### 2020-06-30 2_task_bench.tex
Strip out color from revision.

The Task Bench core library is fully self-validating: {\color{blue} The output of each
task is a tuple $\langle \textrm{row, col}\rangle$ and is unique for a given task
graph. Inputs are verified by checking the expected dependencies
against those received, and an assertion is thrown if validation
fails. These checks ensure that every execution of Task Bench is
correct. Note that the graph representation is concise, making these
checks very inexpensive.} An evaluation of the performance impact of
validation showed it to be less than 3\% at the smallest task
granularities in any Task Bench implementation, with a negligible
effect on overall results.

>>>>

The Task Bench core library is fully self-validating: The output of each
task is a tuple $\langle \textrm{row, col}\rangle$ and is unique for a given task
graph. Inputs are verified by checking the expected dependencies
against those received, and an assertion is thrown if validation
fails. These checks ensure that every execution of Task Bench is
correct. Note that the graph representation is concise, making these
checks very inexpensive. An evaluation of the performance impact of
validation showed it to be less than 3\% at the smallest task
granularities in any Task Bench implementation, with a negligible
effect on overall results.

### 2020-06-30 4_metg.tex
reduce spaces
Strip out color from revision.

The choice of 50\% is a parameter and not fundamental to METG. We use
50\% in our studies to avoid pathologies associated with lower
thresholds (see Section~\ref{subsec:peak-performance-and-efficiency}),
and also because it aligns with what we observe in practice. {\color{blue} For
example, one supercomputer center instructs users applying for
projects to run a strong scaling study and then ``select the most
parallel efficient job size,'' i.e., the largest number of nodes with
a ``ratio of benchmark speed-up vs. linear speed-up above 50\%''
\cite{CSCSReport}, which corresponds to METG(50\%).}

>>>>

The choice of 50\% is a parameter and not fundamental to METG. We use
50\% in our studies to avoid pathologies associated with lower
thresholds (see Section~\ref{subsec:peak-performance-and-efficiency}),
and also because it aligns with what we observe in practice. For
example, one supercomputer center instructs users applying for
projects to run a strong scaling study and then ``select the most
parallel efficient job size,'' i.e., the largest number of nodes with
a ``ratio of benchmark speed-up vs. linear speed-up above 50\%''
\cite{CSCSReport}, which corresponds to METG(50\%).

### 2020-06-30 5_evaluation.tex
Strip out color from revision.

{\color{blue}

@@

}

@@

Not all cores are required to saturate memory bandwidth. {\color{blue}
  This reduces the impact of reserving cores for system use
  (e.g. task-based systems that perform dependence analysis).} Nearly
all systems hit 100\% of peak, unlike the compute-bound case.

@@

We observe that overheads vary by over 5 orders of
magnitude. {\color{blue} The most efficient systems are explicitly parallel
and provide very lightweight mechanisms for parallelism. Task-based systems for HPC tend to be next most efficient, and
provide additional features such as automatic dependency discovery and
data movement. Slower systems tend to be designed primarily for
large-scale data analysis or workflows.} It is worth
remembering that these are \emph{minimum}
effective task granularities. Applications with an average
task granularity of \emph{at least} this value can usually be expected
to execute efficiently. Typical task granularities will
generally be determined by the application domain being
considered. Most notably, for large-scale data analytics workloads, the higher METG values observed for Spark and
TensorFlow are sufficient. In contrast, for high-performance
scientific simulations, task granularities in the millisecond range
are useful, as such applications communicate (e.g., for halo
exchanges) much more frequently.

@@

{\color{blue}
\subsection{Load Imbalance}

@@

The remaining differences are due primarily to
different scheduling behaviors. The execution of 4
simultaneous task graphs only partially mitigates the
load imbalance between tasks. Systems that provide an
additional on-node work stealing capability (such as Chapel with the \texttt{distrib} scheduler) see additional gains in
efficiency at large task granularities. However, the use of
work-stealing queues can also impact throughput at small task
granularities. For example, Chapel's default (non-work-stealing) scheduler outperforms \texttt{distrib} at very small task granularities. We do not consider Charm++ load balancers because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$). We leave analysis of persistent load imbalance to future work.
}

@@

{\color{blue}
\subsection{Validating METG with Time to Solution}

@@

} % color

>>>>

@@

@@

Not all cores are required to saturate memory bandwidth.
  This reduces the impact of reserving cores for system use
  (e.g. task-based systems that perform dependence analysis). Nearly
all systems hit 100\% of peak, unlike the compute-bound case.

@@

We observe that overheads vary by over 5 orders of
magnitude. The most efficient systems are explicitly parallel
and provide very lightweight mechanisms for parallelism. Task-based systems for HPC tend to be next most efficient, and
provide additional features such as automatic dependency discovery and
data movement. Slower systems tend to be designed primarily for
large-scale data analysis or workflows. It is worth
remembering that these are \emph{minimum}
effective task granularities. Applications with an average
task granularity of \emph{at least} this value can usually be expected
to execute efficiently. Typical task granularities will
generally be determined by the application domain being
considered. Most notably, for large-scale data analytics workloads, the higher METG values observed for Spark and
TensorFlow are sufficient. In contrast, for high-performance
scientific simulations, task granularities in the millisecond range
are useful, as such applications communicate (e.g., for halo
exchanges) much more frequently.

@@

\subsection{Load Imbalance}

@@

The remaining differences are due primarily to
different scheduling behaviors. The execution of 4
simultaneous task graphs only partially mitigates the
load imbalance between tasks. Systems that provide an
additional on-node work stealing capability (such as Chapel with the \texttt{distrib} scheduler) see additional gains in
efficiency at large task granularities. However, the use of
work-stealing queues can also impact throughput at small task
granularities. For example, Chapel's default (non-work-stealing) scheduler outperforms \texttt{distrib} at very small task granularities. We do not consider Charm++ load balancers because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$). We leave analysis of persistent load imbalance to future work.

@@

\subsection{Validating METG with Time to Solution}

@@

### 2020-06-30 6_related.tex
Strip out color from revision.

Though smaller than full applications, mini-apps can be challenging to
implement to a level of quality sufficient for conducting comparative
studies between programming systems. {\color{blue} The largest studies
we know of consider at most 7 and 6 programming systems,
respectively~\cite{LULESH13, Deakin19}, and the latter only considers
on-node programming models.} In both cases, the mini-apps under study
require a separate, tuned implementation (in contrast to Task
Bench). Other studies usually lack a comprehensive evaluation, even if
multiple implementations are available:
(a) The PENNANT reference implementation supports
MPI/OpenMP/MPI+OpenMP~\cite{PENNANT}. A follow-up paper presents a
Regent implementation~\cite{Regent15}.
(b) One follow-up paper for the mini-app CoMD describes a Chapel
implementation~\cite{CoMDChapel16} (comparison against reference
only). Additional follow-up papers consider aspects of the reference
implementation only~\cite{CoMDLoadImbalance17,
  CoMDThreadedModels14}.
(c) A report on the Mantevo project~\cite{Mantevo09} describes a number of
mini-apps, but only includes self-comparisons based on reference
implementations.
And (d) A report on MiniAero~\cite{SandiaReportManyTaskRuntimes15} describes
four implementations of the mini-app, but only includes performance
results for three, of which only two can be compared in an
apples-to-apples manner as the last implementation uses structured
rather than unstructured meshes. A follow-up describes another
implementation in Regent~\cite{Regent15} (comparison vs.~reference
only).

>>>>

Though smaller than full applications, mini-apps can be challenging to
implement to a level of quality sufficient for conducting comparative
studies between programming systems. The largest studies
we know of consider at most 7 and 6 programming systems,
respectively~\cite{LULESH13, Deakin19}, and the latter only considers
on-node programming models. In both cases, the mini-apps under study
require a separate, tuned implementation (in contrast to Task
Bench). Other studies usually lack a comprehensive evaluation, even if
multiple implementations are available:
(a) The PENNANT reference implementation supports
MPI/OpenMP/MPI+OpenMP~\cite{PENNANT}. A follow-up paper presents a
Regent implementation~\cite{Regent15}.
(b) One follow-up paper for the mini-app CoMD describes a Chapel
implementation~\cite{CoMDChapel16} (comparison against reference
only). Additional follow-up papers consider aspects of the reference
implementation only~\cite{CoMDLoadImbalance17,
  CoMDThreadedModels14}.
(c) A report on the Mantevo project~\cite{Mantevo09} describes a number of
mini-apps, but only includes self-comparisons based on reference
implementations.
And (d) A report on MiniAero~\cite{SandiaReportManyTaskRuntimes15} describes
four implementations of the mini-app, but only includes performance
results for three, of which only two can be compared in an
apples-to-apples manner as the last implementation uses structured
rather than unstructured meshes. A follow-up describes another
implementation in Regent~\cite{Regent15} (comparison vs.~reference
only).

### 2020-06-30 8_case_study.tex
Tweak wording and space.
Strip out color from revision.

{\color{blue}
Further analysis of Realm efficiency indicated many overheads due to
dynamic graph construction. The Realm developers implemented a new
``subgraph'' API in response to this feedback to amortize the cost of
repeatedly constructing isomorphic task graphs. This API is used in
Task Bench to achieve further speedups in the Realm implementation.

Chapel uses a naive round-robin scheduler by default, which can lead
to unexpected load balance issues because certain language features
(such as remote array assignment) implicitly generate tasks. This
resulted in poor peak performance, even at large task granularities,
which in some cases made it impossible to measure METG (because peak
efficiency did not exceed 50\%). Chapel's other schedulers add
overhead, resulting in higher METGs. We worked around this with
a \lstinline{serial} block.

>>>>

Further analysis of Realm efficiency indicated many overheads due to
dynamic graph construction. The Realm developers implemented a new
``subgraph'' API in response to this feedback to amortize the cost of
repeatedly constructing isomorphic task graphs. This API is used in
Task Bench to achieve further speedups in the Realm implementation.

Chapel uses a naive round-robin scheduler by default, which can lead
to unexpected load balance issues because certain language features
(such as remote array assignment) implicitly generate tasks. This
resulted in poor peak performance, even at large task granularities,
which in some cases made it impossible to measure METG (because peak
efficiency did not exceed 50\%). Chapel's other schedulers add
overhead, resulting in higher METGs. Based on guidance from the Chapel team, we worked around this with
a \lstinline{serial} block.

### 2020-06-30 e11_metg_predict_strong.tex
Fix numbers.
Strip out color from revision.

\begin{table}[t]
\color{blue}
\centering
\small
\begin{tabular}{r | c | c | c | c}
        & \multicolumn{2}{|c|}{Nodes} & \multicolumn{2}{|c}{Time to Solution} \\
Pattern & mean & std.~dev. & mean & std.~dev. \\
\hline
stencil & $1.61\times$ & 0.324 & $1.29\times$ & 0.238 \\
nearest & $1.77\times$ & 0.253 & $1.17\times$ & 0.341 \\
spread  & $1.39\times$ & 0.322 & $1.28\times$ & 0.207 \\
nearest, 4 graphs & $2.15\times$ & 0.635 & $1.31\times$ & 0.777
\end{tabular}

\vspace{-0.20cm}
\caption{\color{blue} Average factor of separation between limit-ideal and limit-actual intersection across the 4 patterns tested in Section~\ref{subsec:scalability}.\label{tab:metg-predict-strong}}
\vspace{-0.5cm}
\end{table}

>>>>

\begin{table}[t]
\centering
\small
\begin{tabular}{r | c | c | c | c}
        & \multicolumn{2}{|c|}{Nodes} & \multicolumn{2}{|c}{Time to Solution} \\
Pattern & mean & std.~dev. & mean & std.~dev. \\
\hline
stencil & $1.61\times$ & 0.324 & $1.29\times$ & 0.238 \\
nearest & $1.77\times$ & 0.253 & $1.24\times$ & 0.310 \\
spread  & $1.39\times$ & 0.322 & $1.30\times$ & 0.171 \\
nearest, 4 graphs & $2.15\times$ & 0.635 & $1.42\times$ & 0.723
\end{tabular}

\vspace{-0.20cm}
\caption{Average factor of separation between limit-ideal and limit-actual intersection across the 4 patterns tested in Section~\ref{subsec:scalability}.\label{tab:metg-predict-strong}}
\vspace{-0.5cm}
\end{table}
